---
title: "Choropleths Using ggplot"
author: "Francisco Rowe ([`@fcorowe`](http://twitter.com/fcorowe))"
date: "`r Sys.Date()`"
output: tint::tintHtml
bibliography: skeleton.bib
link-citations: yes
---

```{r setup, include=FALSE}
library(tint)
# handle spatial data
library(sf)
# manipulate data
library(tidyverse)
library(lubridate)
# create maps
library(tmap)
# nice colour schemes
library(viridis) 
library(viridisLite)
library(RColorBrewer)
library(ggthemes)
library(scico)
library(patchwork)
# invalidate cache when the package version changes
knitr::opts_chunk$set(tidy = FALSE, cache.extra = packageVersion('tint'))
options(htmltools.dir.version = FALSE)
```


# Choropleths

# Data

We will use [internal migration data](https://www.ons.gov.uk/peoplepopulationandcommunity/populationandmigration/migrationwithintheuk/datasets/internalmigrationbyoriginanddestinationlocalauthoritiessexandsingleyearofagedetailedestimatesdataset) from the Office for National Statistics (ONS) from the United Kingdom. 

The original data are organised in a long format structure and are disaggregated by sex and age. Each row captures the count of people moving from an origin to a destination. The spatial units of analysis are local authorities (LA).

```{r}
# clean workspace
rm(list=ls())

# load data
df_long <- read_csv("../data/internal_migration/Detailed_Estimates_2020_LA_2021_Dataset_1.csv")

# id for origins and destinations
orig_la_nm <- as.data.frame(unique(df_long$OutLA))
dest_la_nm <- as.data.frame(unique(df_long$InLA))

head(df_long)
```
We also read our LA boundaries and analyse the structure of the data. We use open data from the ONS's Geography portal. We use the [Local Authority Districts Boundaries (May 2021) UK BFE](https://geoportal.statistics.gov.uk/datasets/ons::local-authority-districts-may-2021-uk-bfe/about)

```{r}
# read shapefile
la_shp <- st_read("../data/Local_Authority_Districts_(May_2021)_UK_BFE_V3/LAD_MAY_2021_UK_BFE_V2.shp")

str(la_shp)
```

## Computing mobility indicators

```{r}
# out-migration
outflows <- df_long %>% 
  group_by(OutLA) %>%
   dplyr::summarise( n = sum(Moves, na.rm = T))

# in-migration
inflows <- df_long %>% 
  group_by(InLA) %>%
   dplyr::summarise( n = sum(Moves, na.rm = T))

# net migration
indicators <- full_join(outflows, 
                        inflows,
                        by = c("OutLA" = "InLA")) %>% 
  mutate_if(is.numeric, ~replace(., is.na(.), 0)) %>% 
  mutate_if(is.numeric, round) %>% 
  rename(
    outflows = n.x,
    inflows = n.y
  ) %>% 
    mutate(
  netflows = ((inflows) - (outflows))
  ) 

```


## Joining spatial data

```{r}
la_shp <- left_join(la_shp, indicators, by = c("LAD21CD" = "OutLA"))
```


## Simplify boundaries

```{r}
# id for country name initial
la_shp$ctry_nm <- substr(la_shp$LAD21CD, 1, 1)
la_shp$ctry_nm <- as.factor(la_shp$ctry_nm)

# simplify boundaries
la_shp_simple <- st_simplify(la_shp, 
                             preserveTopology =T,
                             dTolerance = 1000) # 1km

# ensure geometry is valid
la_shp_simple <- sf::st_make_valid(la_shp_simple)
```


# Mapping

```{r}
la_shp_simple <- la_shp_simple %>% filter(ctry_nm=="E" | ctry_nm=="W")

netm <- ggplot(data = la_shp_simple, aes(fill = netflows)) +
  geom_sf(col = "gray90", size = 0.1) +
  coord_sf() +
  scale_fill_gradient2(low = "darkblue", 
                       mid = "#ffffe0",
                       #mid = scales::muted("purple"), 
                       high = "darkred", 
                       breaks = ) +
  labs(title = "Net migration",
       subtitle = "ONS 2020 estimates",
       fill = "Count") +
  theme_map()

outm <- ggplot(data = la_shp_simple, aes(fill = outflows)) +
  geom_sf(col = "gray90", size = 0.1) +
  coord_sf() +
  scale_fill_gradient2(low = "#ffffe0", 
                       #mid = "blue",
                       #mid = scales::muted("purple"), 
                       high = "darkblue", 
                       breaks = ) +
  labs(title = "Out-migration",
       #subtitle = "ONS 2020 estimates",
       fill = "Count") +
  theme_map() 


inm <- ggplot(data = la_shp_simple, aes(fill = inflows)) +
  geom_sf(col = "gray90", size = 0.1) +
  coord_sf() +
  scale_fill_gradient2(low = "#ffffe0", 
                       #mid = "blue",
                       #mid = scales::muted("purple"), 
                       high = "darkred", 
                       breaks = ) +
  labs(title = "In-migration",
       fill = "Count",
       caption = "Graphic: Francisco Rowe (@fcorowe)") +
  theme_map() 

#png("migration_uk.png", units="in", width=15, height=8, res=300)
netm + outm + inm
#dev.off()
```


```{r eval=FALSE, include=FALSE}
file.edit(
  tint:::template_resources(
    'tint', '..', 'skeleton', 'skeleton.Rmd'
  )
)
```

```{r bib, include=FALSE}
# create a bib file for the R packages used in this document
knitr::write_bib(c('base', 'rmarkdown'), file = 'skeleton.bib')
```