---
title: "Choropleths Using tmap"
author: "Francisco Rowe ([`@fcorowe`](http://twitter.com/fcorowe))"
date: "`r Sys.Date()`"
output: tint::tintHtml
bibliography: skeleton.bib
link-citations: yes
---

```{r setup, include=FALSE}
library(tint)
# handle spatial data
library(sf)
# manipulate data
library(tidyverse)
library(lubridate)
# create maps
library(tmap)
# nice colour schemes
library(viridis) 
library(viridisLite)
library(RColorBrewer)
library(ggthemes)
library(scico)
library(patchwork)
# invalidate cache when the package version changes
knitr::opts_chunk$set(tidy = FALSE, cache.extra = packageVersion('tint'))
options(htmltools.dir.version = FALSE)
```


# Choropleths

# Data

We will use [internal migration data](https://www.ons.gov.uk/peoplepopulationandcommunity/populationandmigration/migrationwithintheuk/datasets/internalmigrationbyoriginanddestinationlocalauthoritiessexandsingleyearofagedetailedestimatesdataset) from the Office for National Statistics (ONS) from the United Kingdom. 

The original data are organised in a long format structure and are disaggregated by sex and age. Each row captures the count of people moving from an origin to a destination. The spatial units of analysis are local authorities (LA).

```{r}
# clean workspace
rm(list=ls())

# load data
df_long <- read_csv("../data/internal_migration/Detailed_Estimates_2020_LA_2021_Dataset_1.csv")

# id for origins and destinations
orig_la_nm <- as.data.frame(unique(df_long$OutLA))
dest_la_nm <- as.data.frame(unique(df_long$InLA))

head(df_long)
```
We also read our LA boundaries and analyse the structure of the data. We use open data from the ONS's Geography portal. We use the [Local Authority Districts Boundaries (May 2021) UK BFE](https://geoportal.statistics.gov.uk/datasets/ons::local-authority-districts-may-2021-uk-bfe/about)

```{r}
# read shapefile
la_shp <- st_read("../data/Local_Authority_Districts_(May_2021)_UK_BFE_V3/LAD_MAY_2021_UK_BFE_V2.shp")

str(la_shp)
```

## Computing mobility indicators

```{r}
# out-migration
outflows <- df_long %>% 
  group_by(OutLA) %>%
   dplyr::summarise( n = sum(Moves, na.rm = T))

# in-migration
inflows <- df_long %>% 
  group_by(InLA) %>%
   dplyr::summarise( n = sum(Moves, na.rm = T))

# net migration
indicators <- full_join(outflows, 
                        inflows,
                        by = c("OutLA" = "InLA")) %>% 
  mutate_if(is.numeric, ~replace(., is.na(.), 0)) %>% 
  mutate_if(is.numeric, round) %>% 
  rename(
    outflows = n.x,
    inflows = n.y
  ) %>% 
    mutate(
  netflows = ((inflows) - (outflows))
  ) 

```


## Joining spatial data

```{r}
la_shp <- left_join(la_shp, indicators, by = c("LAD21CD" = "OutLA"))
```


## Simplify boundaries

```{r}
# id for country name initial
la_shp$ctry_nm <- substr(la_shp$LAD21CD, 1, 1)
la_shp$ctry_nm <- as.factor(la_shp$ctry_nm)

# simplify boundaries
la_shp_simple <- st_simplify(la_shp, 
                             preserveTopology =T,
                             dTolerance = 1000) # 1km

# ensure geometry is valid
la_shp_simple <- sf::st_make_valid(la_shp_simple)
```


# Mapping

```{r}
la_shp_simple <- la_shp_simple %>% filter(ctry_nm=="E" | ctry_nm=="W")

div_pal <- c('#00429d', '#73a2c6', '#ffffe0', '#f4777f', '#93003a')
seq1_pal <- c('#ffffe0', '#4771b2', '#73a2c6', '#a5d5d8', '#00429d')
seq2_pal <- c('#ffffe0', '#ffbcaf', '#f4777f', '#cf3759', '#93003a')

netm <- tm_shape(la_shp_simple) +
  tm_fill(col = "netflows", 
          style = "sd", 
          #palette = "-RdBu", 
          palette = div_pal,
          title = "Net migration", 
          legend.hist = F) +
  tm_borders(lwd = 0) +
  tm_layout(legend.title.size = 1,
          legend.text.size = 0.6,
          #legend.position = c("right","top"),
          legend.bg.color = "white",
          legend.outside = F,
          legend.bg.alpha = 1)

outm <- tm_shape(la_shp_simple) +
  tm_fill(col = "outflows", 
          style = "sd", 
          palette = seq1_pal, 
          title = "Out-migration",
          legend.hist = F) +
  tm_borders(lwd = 0) +
  tm_layout(legend.title.size = 1,
          legend.text.size = 0.6,
          legend.bg.color = "white",
          legend.outside = F,
          legend.bg.alpha = 1)

inm <- tm_shape(la_shp_simple) +
  tm_fill(col = "inflows", 
          style = "sd",
          palette = seq2_pal, 
          title = "In-migration",
          legend.hist = F) +
  tm_borders(lwd = 0) +
  tm_layout(legend.title.size = 1,
          legend.text.size = 0.6,
          legend.bg.color = "white",
          legend.outside = F,
          legend.bg.alpha = 1)

tmap_arrange(netm, outm, inm, 
                     ncol = 3)

#maps <- tmap_arrange(netm, outm, inm, 
#                     ncol = 3)

#tmap_save(tm = maps, filename = "migration_map.png", dpi = 300)
```